### 2022-01-17 RDT-equivalence.rmd
Initial commit

---
title: "RDT - Gold Standard Equivalence"
author: "Callum Arnold"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.dim = c(10, 8), fig.align = "center")
```

# Initial Look {.tabset .tabset-pills}

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(hrbrthemes)
library(kableExtra)
library(lubridate)
library(glue)
library(paletteer)
```

```{r}
theme_set(theme_minimal())
```

```{r}

```