### 2022-01-20 RDT-equivalence.rmd
Update Rmd to look at uncertainty in true cases
Simulate true cases drawn from binomial distribution. Will need to think about how to incorporate uncertainty from testing

---
title: "RDT - Gold Standard Equivalence"
author: "Callum Arnold"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: FALSE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.dim = c(10, 8), fig.align = "center")
```

# Disease Positive -> Test Positive  {.tabset .tabset-pills}

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(hrbrthemes)
library(kableExtra)
library(lubridate)
library(purrr)
library(glue)
library(MetBrewer)
library(future)
library(furrr)
```

```{r}
theme_set(theme_minimal())
```

```{r}
sens_range <- seq(0, 1, 0.05)
spec_range <- seq(0, 1, 0.05)
data <- tidyr::crossing(sens_range, spec_range)
```

## Deterministic

```{r}
prev <- 0.05
tot <- 100

dis_pos <- tot * prev
dis_neg <- tot * (1 - prev)
```

```{r}
test <- pmap_dfr(
    .l = list(sens = data$sens_range, spec = data$spec_range),
    .f = function(sens, spec){
      true_pos <- dis_pos * sens
      true_neg <- dis_neg * spec
      false_pos <- dis_neg * (1 - spec)
      false_neg <- dis_pos * (1 - sens)

      test_pos <- true_pos + false_pos
      test_neg <- true_neg + false_neg

      ppv <- true_pos / test_pos
      npv <- true_neg / test_neg

      tibble(
        sens = sens,
        spec = spec,
        true_pos = true_pos,
        true_neg = true_neg,
        false_pos = false_pos,
        false_neg = false_neg,
        test_pos = test_pos,
        test_neg = test_neg,
        ppv = ppv,
        npv = npv
      )
    }
  )
```

```{r}
ggplot(test, aes(x = sens, y = spec, z = test_pos)) +
  geom_contour_filled() +
  scale_fill_manual(values = met.brewer("OKeeffe2", 10)) +
  labs(
    title = "Number of Positive Tests by Sensitivity and Specificity",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}, Disease Positives = {dis_pos}"),
    x = "Sensitivity",
    y = "Specificity",
    fill = "Number of Positive Tests"
  ) +
  theme(legend.position = "bottom")
```

```{r}
ggplot(test, aes(x = sens, y = spec, z = ppv)) +
  geom_contour_filled() +
  scale_fill_manual(values = met.brewer("Hokusai2", 10)) +
  labs(
    title = "PPV by Sensitivity and Specificity",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}, Disease Positives = {dis_pos}"),
    x = "Sensitivity",
    y = "Specificity",
    fill = "PPV"
  ) +
  theme(legend.position = "bottom")
```

## Uncertainty

```{r}
plan(multisession)
```

```{r}
set.seed(1234)

stoch_dis_pos_neg <- tibble(
  tot = tot,
  pos = rbinom(n = 1e2, size = tot, prob = dis_pos/100)
  ) %>%
  mutate(neg = tot - pos)
```

```{r}
ggplot(stoch_dis_pos_neg, aes(x = pos)) +
  geom_bar() +
  labs(
    title = "Distribution of Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
  )
```

```{r}
stoch_test <- future_pmap_dfr(
  .l = list(
    iter = 1:nrow(stoch_dis_pos_neg),
    dis_pos = stoch_dis_pos_neg$pos,
    dis_neg = stoch_dis_pos_neg$neg
    ),
  .f = function(iter, dis_pos, dis_neg){
    future_pmap_dfr(
      .l = list(
        sens_spec_comb = 1:nrow(data),
        sens = data$sens_range,
        spec = data$spec_range
        ),
      .f = function(sens_spec_comb, sens, spec){
        true_pos <- dis_pos * sens
        true_neg <- dis_neg * spec
        false_pos <- dis_neg * (1 - spec)
        false_neg <- dis_pos * (1 - sens)

        test_pos <- true_pos + false_pos
        test_neg <- true_neg + false_neg

        ppv <- true_pos / test_pos
        npv <- true_neg / test_neg

        tibble(
          iter = iter,
          sens_spec_comb = sens_spec_comb,
          sens = sens,
          spec = spec,
          dis_pos = dis_pos,
          dis_neg = dis_neg,
          true_pos = true_pos,
          true_neg = true_neg,
          false_pos = false_pos,
          false_neg = false_neg,
          test_pos = test_pos,
          test_neg = test_neg,
          ppv = ppv,
          npv = npv
        )
      },
      .options = furrr_options(seed = 12345)
    )
  },
  .options = furrr_options(seed = 12345)
)
```

```{r}
stoch_test %>%
  filter(sens_spec_comb == 1) %>%
  ggplot() +
  geom_bar(aes(x = dis_pos), fill = "black") +
  geom_bar(aes(x = test_pos), fill = "blue") +
  labs(
    title = "Example Distributions of Disease and Test Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}"),
    x = "Number of Positives",
    caption = "Blue = Test Positives, Black = Disease Positives"
  )
```

```{r}
stoch_test2 <- stoch_test %>%
  mutate(
    test_dis_pos_diff = test_pos - dis_pos
  )
```

```{r}
stoch_test2 %>%
  filter(sens_spec_comb == 1) %>%
  ggplot(aes(x = test_dis_pos_diff)) +
  geom_bar() +
  labs(
    title = "Example Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}"),
    x = "Difference in Test Positives and Disease Positives",
    caption = glue("Using sensitivity = {filter(stoch_test2, sens_spec_comb == 1)$sens}, specificity = {filter(stoch_test2, sens_spec_comb == 1)$spec}")
  )
```

```{r}
stoch_test2 %>%
  filter(sens_spec_comb == 100) %>%
  ggplot(aes(x = test_dis_pos_diff)) +
  geom_bar() +
  labs(
    title = "Example Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}"),
    x = "Difference in Test Positives and Disease Positives",
    caption = glue("Using sensitivity = {filter(stoch_test2, sens_spec_comb == 100)$sens}, specificity = {filter(stoch_test2, sens_spec_comb == 100)$spec}")
  )
```

```{r}
stoch_test3 <- future_map_dfr(
  .x = unique(stoch_test2$sens_spec_comb),
  .f = function(.x){
    stoch_test2 %>%
      filter(sens_spec_comb == .x) %>%
      summarize(
        sens_spec_comb = .x,
        sens = unique(sens),
        spec = unique(spec),
        mean_test_dis_pos_diff = mean(test_dis_pos_diff),
        sd_test_dis_pos_diff = sd(test_dis_pos_diff),
        med_test_dis_pos_diff = median(test_dis_pos_diff),
        min_test_dis_pos_diff = min(test_dis_pos_diff),
        max_test_dis_pos_diff = max(test_dis_pos_diff),
        quantile_10 = quantile(test_dis_pos_diff, .10),
        quantile_25 = quantile(test_dis_pos_diff, .25),
        quantile_50 = quantile(test_dis_pos_diff, .50),
        quantile_75 = quantile(test_dis_pos_diff, .75),
        quantile_90 = quantile(test_dis_pos_diff, .90)
        )
  },
  .options = furrr_options(seed = 12345)
)
```

```{r}
stoch_test3 %>%
  ggplot() +
  geom_contour_filled(aes(x = sens, y = spec, z = min_test_dis_pos_diff)) +
  scale_fill_manual(values = met.brewer("Robert", 11)) +
  labs(
    title = "Minimum Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
  )
```

```{r}
stoch_test3 %>%
  ggplot() +
  geom_contour_filled(aes(x = sens, y = spec, z = quantile_10)) +
  scale_fill_manual(values = met.brewer("Robert", 11)) +
  labs(
    title = "10% Quantile Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
  )
```

```{r}
stoch_test3 %>%
  ggplot() +
  geom_contour_filled(aes(x = sens, y = spec, z = mean_test_dis_pos_diff)) +
  scale_fill_manual(values = met.brewer("OKeeffe2", 11)) +
  labs(
    title = "Mean Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
  )
```

```{r}
stoch_test3 %>%
  ggplot() +
  geom_contour_filled(aes(x = sens, y = spec, z = quantile_90)) +
  scale_fill_manual(values = met.brewer("Peru2", 11)) +
  labs(
    title = "90% Quantile Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
  )
```

```{r}
stoch_test3 %>%
  ggplot() +
  geom_contour_filled(aes(x = sens, y = spec, z = max_test_dis_pos_diff)) +
  scale_fill_manual(values = met.brewer("Peru2", 11)) +
  labs(
    title = "Maximum Difference in Test Positives and Disease Positives",
    subtitle = glue("Prevalence = {prev}, Total Tests = {tot}")
  )
```