### 2024-09-26 manuscript/supplemental-appendix.qmd
Rewrite manuscript using quarto

---
title: "The Need to Develop a Holistic Infectious Disease Surveillance System"
header-title: true
authors:
  - name: "Callum R.K. Arnold"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
    corresponding: true
    email: "cfa5228@psu.edu"
  - name: "Alex C. Kong"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Amy K. Winter"
    affiliations:
      - ref: "UGA"
  - name: "William J. Moss"
    affiliations:
      - ref: "Hopkins-IH"
      - ref: "Hopkins-Epi"
  - name: "Bryan N. Patenaude"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Matthew J. Ferrari"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
affiliations:
  - id: "PSU-Bio"
    address: "Department of Biology, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "CIDD"
    address: "Center for Infectious Disease Dynamics, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "Hopkins-IH"
    address: "Department of International Health, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
  - id: "UGA"
    address: "Department of Epidemiology, College of Public Health, University of Georgia, Athens, GA, USA 30602"
  - id: "Hopkins-Epi"
    address: "Department of Epidemiology, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
execute:
  echo: false
  warning: false
  message: false
# bibliography: /Users/cfa5228/Documents/Repos/OutbreakDetection/manuscript/OD.bib
bibliography: OD.bib
bibliographystyle: springer-vancouver
toc: false
format:
  generic-manuscript-typst:
    keep-typ: true
---

# Results