### 2024-11-06 manuscript/ARCHIVE/supplemental-appendix.qmd
Delete old manuscript files, archive qmd, and use CairoMakie to produce SVG plots in particular directory
Going to use Typst with Make to build the manuscript files, so makes
sense to switch to CairoMakie at the same time

---
title: "Individual and Population Level Uncertainty Interact to Determine Performance of Outbreak Detection"
header-title: true
authors:
  - name: "Callum R.K. Arnold"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
    corresponding: true
    email: "cfa5228@psu.edu"
  - name: "Alex C. Kong"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Amy K. Winter"
    affiliations:
      - ref: "UGA"
  - name: "William J. Moss"
    affiliations:
      - ref: "Hopkins-IH"
      - ref: "Hopkins-Epi"
  - name: "Bryan N. Patenaude"
    affiliations:
      - ref: "Hopkins-IH"
  - name: "Matthew J. Ferrari"
    affiliations:
      - ref: "PSU-Bio"
      - ref: "CIDD"
affiliations:
  - id: "PSU-Bio"
    address: "Department of Biology, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "CIDD"
    address: "Center for Infectious Disease Dynamics, Pennsylvania State University, University Park, PA, USA 16802"
  - id: "Hopkins-IH"
    address: "Department of International Health, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
  - id: "UGA"
    address: "Department of Epidemiology, College of Public Health, University of Georgia, Athens, GA, USA 30602"
  - id: "Hopkins-Epi"
    address: "Department of Epidemiology, Johns Hopkins Bloomberg School of Public Health, Baltimore, MD, USA 21205"
execute:
  cache: true
  echo: false
  warning: false
  message: false
# bibliography: /Users/cfa5228/Documents/Repos/OutbreakDetection/manuscript/OD.bib
bibliography: OD.bib
bibliographystyle: springer-vancouver
toc: false
format:
  generic-manuscript-typst:
    keep-typ: true
---

# Results

## Tables

::: {#tbl-optimal-thresholds-accuracy}

```{=typst}
#let accuracy = csv("optimal-thresholds_accuracy.csv")

#table(
  columns: 9,
    fill: (x, y) => {
      if y == 0 {gray}
      if y == 1 {gray}
    },
  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 6, align: center, "Testing Rate"),
  ..accuracy.flatten()
)
```

Mean outbreak detection accuracy of each testing scenario at their specific optimal thresholds, when the average noise incidence is 8 times higher than the average measles incidence. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

:::

::: {#tbl-optimal-thresholds-unavoidable}

```{=typst}
#let optimal_thresholds = csv("optimal-thresholds_unavoidable-cases.csv")

#table(
  columns: 9,
    fill: (x, y) => {
      if y == 0 {gray}
      if y == 1 {gray}
    },
  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 6, align: center, "Testing Rate"),
  ..optimal_thresholds.flatten()
)
```

Mean unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population, when the average noise incidence is 8 times higher than the average measles incidence. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

:::

::: {#tbl-optimal-thresholds-delays}

```{=typst}
#let optimal_thresholds = csv("optimal-thresholds_detection-delays.csv")

#table(
  columns: 9,
    fill: (x, y) => {
      if y == 0 {gray}
      if y == 1 {gray}
    },
  [], table.cell(colspan: 2, align: center, "Test Characteristic"), table.cell(colspan: 6, align: center, "Testing Rate"),
  ..optimal_thresholds.flatten()
)
```

Mean outbreak alert delay (days) of each testing scenario at their specific optimal thresholds, when the average noise incidence is 8 times higher than the average measles incidence. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

:::

## Figures

```{julia plotting-setup}
#| echo: false
include("optimal-thresholds-loading.jl");
include(srcdir("makie-plotting-setup.jl"));
using OutbreakDetection: line_plot
alpha = 0.2
nbanks = 1
legend_rowsize = Makie.Relative(0.05);
clinical_hline = false
```

```{julia}
#| fig-cap: "The difference between the proportion of the time series in outbreak for outbreak detection systems under different testing rates and noise structures. The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate. Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
#| label: fig-outbreak-proportion
line_plot(
    optimal_threshold_characteristics;
    outcome = :proportion_timeseries_in_outbreak,
    ylabel = "Proportion of Time\nSeries In Outbreak",
    alpha = alpha,
    facet_fontsize = 18,
    labelsize = 20,
    show_x_facet_label = true,
    show_y_facet_label = false,
    ylims = (0.5, 0.25),
    force = true,
    save_plot = false,
    clinical_hline = clinical_hline,
    nbanks = nbanks,
    legend_rowsize = legend_rowsize,
)
```

```{julia}
#| fig-cap: "The alert durations of outbreak detection systems under different testing rates and noise structures. The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate. Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
#| label: fig-alert-duration
line_plot(
    optimal_threshold_characteristics;
    outcome = :alert_duration_vec,
    ylabel = "Alert Duration\n(Days)",
    alpha = alpha,
    facet_fontsize = 18,
    labelsize = 20,
    show_x_facet_label = true,
    show_y_facet_label = false,
    ylims = (0, 170),
    force = true,
    save_plot = false,
    clinical_hline = clinical_hline,
    nbanks = nbanks,
    legend_rowsize = legend_rowsize,
)
```

```{julia}
#| fig-cap: "The number of alerts of outbreak detection systems under different testing rates and noise structures. The shaded bands illustrate the 80% central interval, and the solid/dashed lines represent the mean estimate. Solid lines represent tests with 0-day turnaround times, and dashed lines represent tests with result delays."
#| label: fig-num-alerts
line_plot(
    optimal_threshold_characteristics;
    outcome = :nalerts,
    ylabel = "Number of Alerts",
    alpha = alpha,
    facet_fontsize = 18,
    labelsize = 20,
    show_x_facet_label = true,
    show_y_facet_label = false,
    ylims = (0, 350),
    force = true,
    save_plot = false,
    clinical_hline = clinical_hline,
    nbanks = nbanks,
    legend_rowsize = legend_rowsize,
)
```